---
title: "Staphylococcus epidermidis Analyses"
author: "Stephanie McMahon"
date: "2025-05-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The only things done in R for the S. epidermidis paper were the heatmaps for ARG and Virulence factors
    #Multiple heatmaps mde for the following
      -ARG for AML and Non-AML Isolates
      -ARG for only AML Isolates
      -Virulence for AML and Non-AML Isolates
      -Vurlence for only AML Isolates

#Heatmap for ARG for AML and Non-AML Isolates
```{r ARGs for All Isolates, message=FALSE, results='hide', warning=FALSE}
BiocManager::install("pheatmap")
BiocManager::install("RColorBrewer")
setwd("~/Desktop")
library(pheatmap)
library(RColorBrewer)
library(readxl)


df <- read_excel("Staph_Heatmap_Master_Sheet.xlsx", sheet = "ARG")
dfdata =df[-1]
hm<- data.matrix(dfdata[4:46,])

row.names(hm) = unlist(df[4:46,1])

index <- data.frame(df[1:3,])
id <- index[,2:115]
rownames(id) <- index$ID
colnames(id) <- colnames(hm)
idk <- as.data.frame(t(id))

colors <- list("Origin"=c("AML"="violetred", "Non-AML"="purple4"), "Classification"=c("Colonization"="royalblue3", "Infection"="thistle4"),"Location"=c("Stool"="sienna4", "Blood"="red2", "Skin"="peachpuff1", "Central Venous Catheter"="cyan", "Pertoneal Cavity"="purple", "Cerebrospinal Fluid"="steelblue", "Urine"="goldenrod1", "Nasal Swab"="darkorange3", "Ear Swab"="slateblue3", "Tissue"="seagreen1"))

pheatmap(hm, color = c("white", "forestgreen"), annotation_col = idk, annotation_colors = colors, legend = FALSE, border_color = NA)
```
#ARG for only AML Isolates
```{r ARGs for AML Isolates ONLY, message=FALSE, results='hide'}
df <- read_excel("Staph_Heatmap_Master_Sheet.xlsx", sheet = "Lab ARG")
dfdata =df[-1]
hm<- data.matrix(dfdata[3:46,])

row.names(hm) = unlist(df[3:46,1])


index <- data.frame(df[1:2,])
id <- index[,2:45]
rownames(id) <- index$ID
colnames(id) <- colnames(hm)
idk <- as.data.frame(t(id))

colors2 <- list("Classification"=c("Colonization"="royalblue3", "Infection"="thistle4"),"Location"=c("Stool"="sienna4", "Blood"="red2", "Tissue"="seagreen1", "Urine"="goldenrod1"))

pheatmap(hm, color = c("white", "forestgreen"), annotation_col = idk, annotation_colors = colors2, legend = FALSE, border_color = NA)
```

#Virulence for AML and Non-AML Isolates
```{r Virulence Genes for All Isolates, message=FALSE, results='hide'}
df <- read_excel("Staph_Heatmap_Master_Sheet.xlsx", sheet = "Virulence")
dfdata =df[-1]
hm<- data.matrix(dfdata[4:8,])

row.names(hm) = unlist(df[4:8,1])

index <- data.frame(df[1:3,])
id <- index[,2:115]
rownames(id) <- index$ID
colnames(id) <- colnames(hm)
idk <- as.data.frame(t(id))

colors <- list("Origin"=c("AML"="violetred", "Non-AML"="purple4"), "Classification"=c("Colonization"="royalblue3", "Infection"="thistle4"),"Location"=c("Stool"="sienna4", "Blood"="red2", "Skin"="peachpuff1", "Central Venous Catheter"="cyan", "Pertoneal Cavity"="purple", "Cerebrospinal Fluid"="steelblue", "Urine"="goldenrod1", "Nasal Swab"="darkorange3", "Ear Swab"="slateblue3", "Tissue"="seagreen1"))
pheatmap(hm, color = c("white", "forestgreen"), annotation_col = idk, annotation_colors = colors, legend = FALSE, border_color = NA)
```

#Virulence for only AML Isolates
```{r Virulence Genes for AML Isolates ONLY, message=FALSE, results='hide'}
df <- read_excel("Staph_Heatmap_Master_Sheet.xlsx", sheet = "Lab Virulence")
dfdata =df[-1]
hm<- data.matrix(dfdata[3:8,])

row.names(hm) = unlist(df[3:8,1])


index <- data.frame(df[1:2,])
id <- index[,2:45]
rownames(id) <- index$ID
colnames(id) <- colnames(hm)
idk <- as.data.frame(t(id))

colors2 <- list("Classification"=c("Colonization"="royalblue3", "Infection"="thistle4"),"Location"=c("Stool"="sienna4", "Blood"="red2", "Tissue"="seagreen1", "Urine"="goldenrod1"))


pheatmap(hm, color = c("white", "forestgreen"), annotation_col = idk, annotation_colors = colors2, legend = FALSE, border_color = NA)
```









